Interpolation and Extrapolation Performance Measurement of Analytical and ANN-Based Flow Laws for Hot Deformation Behavior of Medium Carbon Steel

### 2023-02-06 Article.tex
Small update of figure captions

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{In the present work, a critical analysis of the most commonly used analytical models and recently introduced ANN-based models is performed to evaluate their predictive accuracy within and outside the experimental interval used to generate them.
The high temperature deformation behavior of a medium carbon steel was studied over a wide range of strains, strain rates, and temperatures using hot compression tests on a Gleeble-3800.
The experimental flow curves were modeled using Johnson-Cook, Modified-Zerilli-Armstrong, Hansel-Spittel, Arrhenius, and PTM models, as well as an ANN model.
The mean absolute relative error and root mean square error values were used to quantify the predictive accuracy of the models analyzed.
The results indicated that the Johnson-Cook and Modified-Zerilli-Armstrong models had a significant error, while the Hansel-Spittel, PTM, and Arrhenius models were able to predict the behavior of this alloy.
The ANN model showed excellent agreement, with less than a 0.62\% error, between the predicted and experimental flow curves.
To validate the performance, the ability to interpolate and extrapolate the experimental data was also tested.
The Hansel-Spittel, PTM, and Arrhenius models showed good interpolation and extrapolation capabilities.
However, the ANN model was the most powerful of all the models.}

@@

The values predicted by the Johnson-Cook flow law (solid line) and the experimental values (dots) are compared in Figure \ref{fig:CompExp-JC-6}.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]{Figures/CompExp-JC-6}
\caption{Comparison between the experimental flow stresses and those predicted by the JC model}
\label{fig:CompExp-JC-6}
\end{figure}
The JC model is unable to describe the evolution of the flow stress of medium carbon for all strain values, and the predicted stresses $\sigma^y$ show a monotonically decreasing trend for several strain rate values.
The discrepancy between the predicted and experimental values is large for low strains and sometimes acceptable for high strains.
As expected, the mathematical formulation of the Johnson-Cook flow law is unable to represent the stress drop at low strain rates, with the JC model increasing only monotonically with the strain.
Since most of the parameters are identified at low strain rates ($A$, $B$ and $n$), this results in a very poor fit of the model to the experimental data.

@@

The MZA model, which is the modified form of the ZA model, like the JC model presented earlier, is one of the most widely used models implemented in many FEA codes such as the Abaqus software.
The difference between the MZA model and JC model is related to the consideration of the three physical parameters to describe the reality observed from experiments.
In the JC model, the parameters are considered separately, while in MZA, they are considered simultaneously, and for that formulation, MZA is preferred to the JC model \cite{zhu2022thermal}.
However, the original form of the ZA model has some limitations due to the fact that it is considered as a two terms function (thermal and athermal functions), and to improve the formulation, Samantaray \eal \cite{Samantaray-2009} proposed a modified form given by the following equation:
\begin{equation}
\label{eq:MZA-model}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + \left(C_{5}+C_{6}\left(T-T_0\right)\right)\ln\left( \frac{\mdot\varepsilon}{\mdot{\varepsilon}_0}\right)\right],
\end{equation}
where the $7$ coefficients $C_i$ and $n$ are the parameters of the model to be identified for a given material.
To get the parameters of the MZA model, we apply the method proposed by \cite{Samantaray-2009}, and the parameters are summarized in Table \ref{tab:MZA}, while their predictions are plotted in Figure \ref{fig:CompExp-MZA-6}.
\begin{table}[h!]
\centering{}
\caption{Parameter constants of Modified Zerilli-Armstrong model}
\begin{tabular}{ccccccc}
	\toprule
	\boldmath{$C_1$} & \boldmath{$C_2$} &    \boldmath{$C_3$}    &    \boldmath{$C_4$}    & \boldmath{$C_5$} &   \boldmath{$C_6$}    & \boldmath{$n$} \\ \midrule
	   $13.5143$     &    $21.2591$     & $4.7902\times 10^{-3}$ & $1.4895\times 10^{-4}$ &     $0.1389$     & $1.495\times 10^{-4}$ &    $0.0621$    \\ \bottomrule
\end{tabular}
\label{tab:MZA}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-MZA-6}
\caption{Comparison between the experimental flow stresses and those predicted by the MZA model}
\label{fig:CompExp-MZA-6}
\end{figure}
It can be seen from this figure that the MZA model as well as the JC model is not able to faithfully reproduce the experimental data at low strain rates, but is much better for high strain rates.
The deviation between the predicted values and the experimental values is large because this model has a problem correctly describing the softening in its formulation.
For the MZA model, $\MARE=21.20~\%$ and $\RMSE=19.57~\MPa$, showing an overall worse performance than the JC model, as reported in \cite{TizeMha-2022}.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-HS-6}
\caption{Comparison between the experimental flow stresses and those predicted by the HS model}
\label{fig:CompExp-HS-6}
\end{figure}

@@

Figure \ref{fig:CompExp-AR-6} shows a comparison of the values predicted by the Arrhenius model and the experimental values.
The difference between the experimental and predicted values is small.
However for the strain rate $\mdot\varepsilon = 0.01~\ps$ and for the two low temperature values, the AR model is unable to predict the softening.
For the AR model, $\MARE=3.56~\%$ and $\RMSE=2.18~\MPa$.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-AR-6}
\caption{Comparison between the experimental flow stresses and those predicted by the AR model}
\label{fig:CompExp-AR-6}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-PTM-6}
\caption{Comparison between the experimental flow stresses and those predicted the PTM model}
\label{fig:CompExp-PTM-6}
\end{figure}

@@

As for the above considered analytical models, Figure \ref{fig:CompExp-3-15-7-1-sigmoid} shows a comparison between the flow stresses predicted by the ANN model and the data measured during the hot compression test.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-3-15-7-1-sigmoid}
\caption{Comparison between the experimental flow stresses and those predicted by the 3-15-7-1-sigmoid ANN model}
\label{fig:CompExp-3-15-7-1-sigmoid}
\end{figure}
The experimental data and the ANN prediction correlate very well over the entire range of data and the predicted data can track the hardening and softening regions of the hot deformed material well.
For the ANN model, $\MARE=0.62~\%$ and $\RMSE=0.38~\MPa$ which is excellent.
This model can be used to simulate the hot deformation of this type of alloy with much greater reliability to the actual material behavior than the analytical models presented in the above sections.

@@

Figure \ref{fig:CompInt} shows a comparison, for strain rate $\mdot\varepsilon=1~\ps$, of the flow stresses $\sigma^y$ calculated by the models (as a line) and the experimental results (as dots) for the 5 analytical models and the neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompInt}
\caption{Comparison between the experimental (dot) and predicted flow stresses $\sigma^y$ (lines) for $\mdot\varepsilon=1~\ps$}
\label{fig:CompInt}
\end{figure}
Table \ref{tab:IntVal} shows the $\MARE$ and $\RMSE$ deviations between the different models and the experimental data calculated either for the $5$ identified strain rates (lines referred as id. $\mdot\varepsilon$), the $6$ strain rates (lines referred as all $\mdot\varepsilon$) or only on the strain rate $\mdot\varepsilon=1~\ps$.
\begin{table}[h!]
\centering{}
\caption{Accuracy coefficients of interpolation for all models with $\mdot\varepsilon=1~\ps$}
\begin{tabular}{llcccccc}
	\toprule
	\textbf{Strain rate} & \textbf{Coefficients} & \textbf{JC} & \textbf{MZA} & \textbf{HS} & \textbf{AR} & \textbf{PTM} & \textbf{ANN} \\ \midrule
	\mr{2}{id. $\mdot\varepsilon$} & $\MARE(\%)$ & $13.79$ & $20.22$ & $8.57$ & $3.96$ & $5.10$ & $0.70$ \\
	 & $\RMSE(\MPa)$ & $11.83$ & $19.00$ & $3.97$ & $2.29$ & $4.73$ & $0.38$ \\ \midrule
	\mr{2}{$\mdot\varepsilon=1~\ps$} & $\MARE(\%)$ & $14.87$ & $27.79$ & $3.74$ & $1.46$ & $3.16$ & $2.47$ \\
	 & $\RMSE(\MPa)$ & $12.02$ & $23.67$ & $3.10$ & $1.46$ & $2.65$ & $2.77$ \\ \midrule
	\mr{2}{all $\mdot\varepsilon$} & $\MARE(\%)$ & $14.42$ & $28.90$ & $7.55$ & $3.45$ & $4.90$ & $0.96$ \\
	 & $\RMSE(\MPa)$ & $11.86$ & $19.86$ & $3.84$ & $2.17$ & $4.45$ & $1.18$ \\ \bottomrule
\end{tabular}
\label{tab:IntVal}
\end{table}

@@

In this new identification configuration, Figure \ref{fig:CompExt} shows a comparison, for strain rate $\mdot\varepsilon=5~\ps$, of the flow stresses $\sigma^y$ computed by the models (as a line) and the experimental results (as dots) for the 5 analytical models and the neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExt}
\caption{Comparison between the experimental (dot) and predicted flow stresses $\sigma^y$ (lines) for $\mdot\varepsilon=5~\ps$}
\label{fig:CompExt}
\end{figure}
Table \ref{tab:ExtVal} shows the $\MARE$ and $\RMSE$ deviations between the different models and the experimental data calculated either for the $5$ identified strain rates (lines referred as id. $\mdot\varepsilon$), the $6$ strain rates (lines referred as all $\mdot\varepsilon$) or only on the strain rate $\mdot\varepsilon=5~\ps$.
\begin{table}[h!]
\centering{}
\caption{Accuracy coefficients of extrapolation for all models with $\mdot\varepsilon=5~\ps$}
\begin{tabular}{llcccccc}
	\toprule
	\textbf{Strain rate}             & \textbf{Coefficients} & \textbf{JC} & \textbf{MZA} & \textbf{HS} & \textbf{AR} & \textbf{PTM} & \textbf{ANN} \\ \midrule
	\mr{2}{id. $\mdot\varepsilon$}   & $\MARE(\%)$           &   $13.09$   &   $19.09$    &   $7.30$    &   $4.03$    &    $4.34$    &    $0.61$    \\
	                                 & $\RMSE(\MPa)$         &   $9.86$    &   $16.26$    &   $3.36$    &   $2.32$    &    $3.63$    &    $0.32$    \\ \midrule
	\mr{2}{$\mdot\varepsilon=5~\ps$} & $\MARE(\%)$           &   $20.16$   &   $24.27$    &   $7.73$    &   $3.53$    &   $11.46$    &    $3.87$    \\
	                                 & $\RMSE(\MPa)$         &   $20.73$   &   $25.95$    &   $7.83$    &   $4.02$    &   $12.91$    &    $5.84$    \\ \midrule
	\mr{2}{all $\mdot\varepsilon$}   & $\MARE(\%)$           &   $15.12$   &   $25.87$    &   $7.12$    &   $3.86$    &    $5.34$    &    $1.09$    \\
	                                 & $\RMSE(\MPa)$         &   $12.36$   &   $18.24$    &   $4.43$    &   $2.68$    &    $6.23$    &    $2.40$    \\ 
	                                 \bottomrule
\end{tabular}
\label{tab:ExtVal}
\end{table}

>>>>

% Abstract (Do not insert blank lines, i.e. \\) 
\abstract{In the present work, a critical analysis of the most commonly used analytical models and recently introduced ANN-based models is performed to evaluate their predictive accuracy within and outside the experimental interval used to generate them.
The high temperature deformation behavior of a medium carbon steel was studied over a wide range of strains, strain rates, and temperatures using hot compression tests on a Gleeble-3800.
The experimental flow curves were modeled using Johnson-Cook, Modified-Zerilli-Armstrong, Hansel-Spittel, Arrhenius, and PTM models, as well as an ANN model.
The mean absolute relative error and root-mean-square error values were used to quantify the predictive accuracy of the models analyzed.
The results indicated that the Johnson-Cook and Modified-Zerilli-Armstrong models had a significant error, while the Hansel-Spittel, PTM, and Arrhenius models were able to predict the behavior of this alloy.
The ANN model showed excellent agreement between the predicted and experimental flow curves, with an error of less than 0.62\%.
To validate the performance, the ability to interpolate and extrapolate the experimental data was also tested.
The Hansel-Spittel, PTM, and Arrhenius models showed good interpolation and extrapolation capabilities.
However, the ANN model was the most powerful of all the models.}

@@

The values predicted by the Johnson-Cook flow law (solid line) and the experimental values (dots) are compared in Figure \ref{fig:CompExp-JC-6}.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]{Figures/CompExp-JC-6}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ by the Johnson-Cook model}
\label{fig:CompExp-JC-6}
\end{figure}
The JC model is unable to describe the evolution of the flow stress of medium carbon for all strain values, and the predicted stresses $\sigma^y$ show a monotonically decreasing trend for several strain rate values.
The discrepancy between the predicted and experimental values is large for low strains and sometimes acceptable for high strains.
As expected, the mathematical formulation of the Johnson-Cook flow law is unable to represent the stress drop at low strain rates, with the JC model increasing only monotonically with the strain.
Since most of the parameters are identified at low strain rates ($A$, $B$ and $n$), this results in a very poor fit of the model to the experimental data.

@@

The MZA model, which is the modified form of the ZA model, like the JC model presented earlier, is one of the most widely used models implemented in many FEA codes such as the Abaqus software.
The difference between the MZA model and JC model is related to the consideration of the three physical parameters to describe the reality observed from experiments.
In the JC model, the parameters are considered separately, while in MZA, they are considered simultaneously, and for that formulation, MZA is preferred to the JC model \cite{zhu2022thermal}.
However, the original form of the ZA model has some limitations due to the fact that it is considered as a two terms function (thermal and athermal functions), and to improve the formulation, Samantaray \eal \cite{Samantaray-2009} proposed a modified form given by the following equation:
\begin{equation}
\label{eq:MZA-model}
\sigma^y = \left(C_{1}+C_{2}\varepsilon^{p^n}\right) \exp\left[-\left(C_{3}+C_{4}\varepsilon^p\right)\left(T-T_0\right) + \left(C_{5}+C_{6}\left(T-T_0\right)\right)\ln\left( \frac{\mdot\varepsilon}{\mdot{\varepsilon}_0}\right)\right],
\end{equation}
where the $7$ coefficients $C_i$ and $n$ are the parameters of the model to be identified for a given material.
To get the parameters of the MZA model, we apply the method proposed by \cite{Samantaray-2009}, and the parameters are summarized in Table \ref{tab:MZA}, while their predictions are plotted in Figure \ref{fig:CompExp-MZA-6}.
\begin{table}[h!]
\centering{}
\caption{Parameter constants of Modified Zerilli-Armstrong model}
\begin{tabular}{ccccccc}
	\toprule
	\boldmath{$C_1$} & \boldmath{$C_2$} &    \boldmath{$C_3$}    &    \boldmath{$C_4$}    & \boldmath{$C_5$} &   \boldmath{$C_6$}    & \boldmath{$n$} \\ \midrule
	   $13.5143$     &    $21.2591$     & $4.7902\times 10^{-3}$ & $1.4895\times 10^{-4}$ &     $0.1389$     & $1.495\times 10^{-4}$ &    $0.0621$    \\ \bottomrule
\end{tabular}
\label{tab:MZA}
\end{table}
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-MZA-6}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ by the Modified Zerilli-Armstrong model}
\label{fig:CompExp-MZA-6}
\end{figure}
It can be seen from this figure that the MZA model as well as the JC model is not able to faithfully reproduce the experimental data at low strain rates, but is much better for high strain rates.
The deviation between the predicted values and the experimental values is large because this model has a problem correctly describing the softening in its formulation.
For the MZA model, $\MARE=21.20~\%$ and $\RMSE=19.57~\MPa$, showing an overall worse performance than the JC model, as reported in \cite{TizeMha-2022}.

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-HS-6}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ by the Hansel-Spittel model}
\label{fig:CompExp-HS-6}
\end{figure}

@@

Figure \ref{fig:CompExp-AR-6} shows a comparison of the values predicted by the Arrhenius model and the experimental values.
The difference between the experimental and predicted values is small.
However for the strain rate $\mdot\varepsilon = 0.01~\ps$ and for the two low temperature values, the AR model is unable to predict the softening.
For the AR model, $\MARE=3.56~\%$ and $\RMSE=2.18~\MPa$.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-AR-6}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ by the Arrhenius model}
\label{fig:CompExp-AR-6}
\end{figure}

@@

\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-PTM-6}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ by the PTM model}
\label{fig:CompExp-PTM-6}
\end{figure}

@@

As for the above considered analytical models, Figure \ref{fig:CompExp-3-15-7-1-sigmoid} shows a comparison between the flow stresses predicted by the ANN model and the data measured during the hot compression test.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExp-3-15-7-1-sigmoid}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ by the 3-15-7-1-sigmoid ANN model}
\label{fig:CompExp-3-15-7-1-sigmoid}
\end{figure}
The experimental data and the ANN prediction correlate very well over the entire range of data and the predicted data can track the hardening and softening regions of the hot deformed material well.
For the ANN model, $\MARE=0.62~\%$ and $\RMSE=0.38~\MPa$ which is excellent.
This model can be used to simulate the hot deformation of this type of alloy with much greater reliability to the actual material behavior than the analytical models presented in the above sections.

@@

Figure \ref{fig:CompInt} shows a comparison, for strain rate $\mdot\varepsilon=1~\ps$, of the flow stresses $\sigma^y$ calculated by the models (as a line) and the experimental results (as dots) for the 5 analytical models and the neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompInt}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ for $\mdot\varepsilon=1~\ps$}
\label{fig:CompInt}
\end{figure}
Table \ref{tab:IntVal} shows the $\MARE$ and $\RMSE$ deviations between the different models and the experimental data calculated either for the $5$ identified strain rates (lines referred as id. $\mdot\varepsilon$), the $6$ strain rates (lines referred as all $\mdot\varepsilon$) or only on the strain rate $\mdot\varepsilon=1~\ps$.
\begin{table}[h!]
\centering{}
\caption{Accuracy coefficients of interpolation for all models with $\mdot\varepsilon=1~\ps$}
\begin{tabular}{llcccccc}
	\toprule
	\textbf{Strain rate} & \textbf{Coefficients} & \textbf{JC} & \textbf{MZA} & \textbf{HS} & \textbf{AR} & \textbf{PTM} & \textbf{ANN} \\ \midrule
	\mr{2}{id. $\mdot\varepsilon$} & $\MARE(\%)$ & $13.79$ & $20.22$ & $8.57$ & $3.96$ & $5.10$ & $0.70$ \\
	 & $\RMSE(\MPa)$ & $11.83$ & $19.00$ & $3.97$ & $2.29$ & $4.73$ & $0.38$ \\ \midrule
	\mr{2}{$\mdot\varepsilon=1~\ps$} & $\MARE(\%)$ & $14.87$ & $27.79$ & $3.74$ & $1.46$ & $3.16$ & $2.47$ \\
	 & $\RMSE(\MPa)$ & $12.02$ & $23.67$ & $3.10$ & $1.46$ & $2.65$ & $2.77$ \\ \midrule
	\mr{2}{all $\mdot\varepsilon$} & $\MARE(\%)$ & $14.42$ & $28.90$ & $7.55$ & $3.45$ & $4.90$ & $0.96$ \\
	 & $\RMSE(\MPa)$ & $11.86$ & $19.86$ & $3.84$ & $2.17$ & $4.45$ & $1.18$ \\ \bottomrule
\end{tabular}
\label{tab:IntVal}
\end{table}

@@

In this new identification configuration, Figure \ref{fig:CompExt} shows a comparison, for strain rate $\mdot\varepsilon=5~\ps$, of the flow stresses $\sigma^y$ computed by the models (as a line) and the experimental results (as dots) for the 5 analytical models and the neural network.
\begin{figure}[!ht]
\centering
\includegraphics[width=\columnwidth]
{Figures/CompExt}
\caption{Comparison between the experimental (dots) and predicted (lines) flow stresses $\sigma^y$ for $\mdot\varepsilon=5~\ps$}
\label{fig:CompExt}
\end{figure}
Table \ref{tab:ExtVal} shows the $\MARE$ and $\RMSE$ deviations between the different models and the experimental data calculated either for the $5$ identified strain rates (lines referred as id. $\mdot\varepsilon$), the $6$ strain rates (lines referred as all $\mdot\varepsilon$) or only on the strain rate $\mdot\varepsilon=5~\ps$.
\begin{table}[h!]
\centering{}
\caption{Accuracy coefficients of extrapolation for all models with $\mdot\varepsilon=5~\ps$}
\begin{tabular}{llcccccc}
	\toprule
	\textbf{Strain rate}             & \textbf{Coefficients} & \textbf{JC} & \textbf{MZA} & \textbf{HS} & \textbf{AR} & \textbf{PTM} & \textbf{ANN} \\ \midrule
	\mr{2}{id. $\mdot\varepsilon$}   & $\MARE(\%)$           &   $13.09$   &   $19.09$    &   $7.30$    &   $4.03$    &    $4.34$    &    $0.61$    \\
	                                 & $\RMSE(\MPa)$         &   $9.86$    &   $16.26$    &   $3.36$    &   $2.32$    &    $3.63$    &    $0.32$    \\ \midrule
	\mr{2}{$\mdot\varepsilon=5~\ps$} & $\MARE(\%)$           &   $20.16$   &   $24.27$    &   $7.73$    &   $3.53$    &   $11.46$    &    $3.87$    \\
	                                 & $\RMSE(\MPa)$         &   $20.73$   &   $25.95$    &   $7.83$    &   $4.02$    &   $12.91$    &    $5.84$    \\ \midrule
	\mr{2}{all $\mdot\varepsilon$}   & $\MARE(\%)$           &   $15.12$   &   $25.87$    &   $7.12$    &   $3.86$    &    $5.34$    &    $1.09$    \\
	                                 & $\RMSE(\MPa)$         &   $12.36$   &   $18.24$    &   $4.43$    &   $2.68$    &    $6.23$    &    $2.40$    \\ 
	                                 \bottomrule
\end{tabular}
\label{tab:ExtVal}
\end{table}
